Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-11-07 papers/thesis-kirstie/thesis.tex
edits

The goal of this project is to generate computationally temperature-density 
and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
model fluid using classical Density Functional Theory (cDFT). 
To do this we need to find the Helmholtz free energies that the fluid
would have in a liquid state, and in a crystalline solid state 
at a given temperature and density. 
In general, the Helmholtz free energy, which consists of an ideal part 
and an excess part, can be written as a functional of the number density 
profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}

Classical Density Functional Theory is used to find the equilibrium 
Helmholtz free energy, and Soft Fundamental Measure Theory (SFMT) will 
be used to form the functional for the excess Helmholtz free energy. 
According to classical Density Functional Theory the equilbrium 
Helmholtz free energy of the system can be found by varying the number 
density profile $n(\vec{r})$ until the free energy is minimized. 
SFMT makes the task of forming an expression
for the excess Helmholtz free energy of a particular model fluid 
much easier since the interaction that each sphere, or atom, 
has with every other sphere does not have to be taken explicitly into account.
Results will be compared to those from Monte-Carlo simulations.

@@

Real fluids and their behavior are very complex from a physicist's point 
of view. Although a water molecule, for example, is composed of only three
atoms, the ever-changing redistribution of charges within the molecules, 
and their interaction with charges from other molecules make it difficult 
to study a fluid composed of water molecules directly. The task becomes 
even more daunting considering the complicated molecular motion of each 
water molecule, and the sheer number of molecules that must be taken 
into account.

The problem of understanding the behavoir of real liquids is better addressed 
by using models which reduce a complex picture to a simple one. Modifications, 
or perturbations, can then be incorporated into the model making it 
progressively more complex until it mimics the behavior of a real 
liquid as closely as possible.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{plot_HardSphere_Potential.pdf}
    \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity 
    for $r\leq{2R}$ which is approximately 1.225 in this example.}
    \label{fig:HardSphere_potential}
  \end{figure}

@@

The repulsive portion of the Leonard-Jones potential forms the potential for a 
Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship}, 
shown in Figure \ref{fig:WCA_potential}. 
The WCA potential is described by the equation: 
\begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation} 
As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases 
from infinity as the distance between the centers of the atoms r increases, 
and becomes, and remains, essentially zero when 
r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. 
As with the Lenard-Jones potential, the repulsive WCA potential increases 
rapidly and approches infinity as the two spheres begin to overlap. This 
is a more realistic repulsive potential than that of the Hard Sphere model 
which goes instantly to infinity when the two spheres make contact (at r=2R), 
and is zero otherwise. Because the WCA spheres have no hard boundary in 
their interatomic potential, they are called ``soft spheres'' and their 
gradually fading potential makes them slightly ``squishy''.

@@

%For each particular temperature, the pressure of the liquid and the 
%solid are the same at the point of transistion. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{P-T_Diagram.png}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
  \end{figure}
The pressure can be related to the Helmholtz free energy F, which is minimized 
for a system at fixed temperature, volume and number of particles as the 
entropy of the system and its surroundings is maximized. The negative of 
the partial derivative of the Helmholtz free energy with respect to volume V 
at a fixed temperature T and number of particles N equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the 
number of particles, the partial derivative of the Helmholtz free energy 
with respect to volume becomes the partial derivative of the Helmholtz 
free energy per atom with respect to the inverse number density. 
\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.

%In a process called Maxwell's Double Tangent construction, two curves are 
%used to find the densities of the liquid and the solid at the point of transistion. 
%Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for 
%the solid both at a fixed temperature. Since the pressure of the liquid 
%is the same as the pressure of the solid at the point of transistion, the 
%density of the liquid and the density of the solid at the point of freezing 
%can be found from identifying the inverse densties at which the tangents 
%to each curve have the same slope.%
Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
and one for the solid at a fixed temperature. Since the pressure of the 
liquid is the same as the pressure of the solid at the point of transistion, 
the density of the liquid and the density of the solid at the point of 
freezing can be found from identifying the inverse densties at which the 
tangents to each curve have the same slope. This process is called 
Maxwell's Double Tangent Construction.

@@

%\begin{displaymath}{F=-PV+G}\end{displaymath} 
\begin{displaymath}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{displaymath} 
\begin{displaymath}{f=-P\frac{1}{n}+g}\end{displaymath}
This equation has the form of a line:
\begin{displaymath}{y=mx+y_o}\end{displaymath}
%\begin{displaymath}{G_1=G_2}\end{displaymath} 
%\begin{displaymath}{F_1+PV_1=F_2+PV_2}\end{displaymath} 
%\begin{displaymath}{\frac{F_1}{N}+P\frac{V_1}{N}=\frac{F_2}{N}+P\frac{V_2}{N}}\end{displaymath} 
and describes the red line shown in Figure \ref{fig:MaxwellDT} with the 
pressure given by the negative of the slope of the line, and the Gibb's 
free energy per atom given by the y-intercept. At the point of phase transition:
\begin{displaymath}{g_L=g_S}\end{displaymath} 
\begin{displaymath}{f_L+P\frac{1}{n_L}=f_S+P\frac{1}{n_S}}\end{displaymath}

The pressure at the transistion point can be obtained from the point of 
intersection that occurs when the Gibbs free energy per atom is plotted 
against the pressure for both the liquid and the solid at fixed temperature, 
as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from 
the slope and cooresponding y-intercept data collected from lines tangent 
along the curves on the the Helmholtz free energy per atom verses inverse 
density plots for both the liquid and the solid. 
\begin{figure}[h!]
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
     %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

@@

\begin{displaymath}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{displaymath}    \begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$. 
%\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    
%$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%  
The function $f_{ij}$ is called a Mayer function, and is given by
\begin{equation}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
Multiplying the Mayer function terms, the excess free energy can be written as
\begin{equation}\label{eq:Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)  }}\end{equation}

@@

The weighted densities can now be expressed as
\begin{equation}{\color{red} CHECK!~~\color{black}n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}}\end{equation} 
with weight functions
\begin{equation}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
\begin{equation}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}

%\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{R}}}\end{equation}
\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
%\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{R}^2}}\end{equation}
\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
where the weight functions $w_{2}$ and $w_{3}$ have the relation
\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}

The weight functions can be related to the mayer function \begin{equation}{f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1}\end{equation} where r is the distance between the centers of two hard spheres. For a hard sphere fluid the interatomic potential $V(r)$ is given by \begin{displaymath}{V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.}\end{displaymath}
and so the mayer function for a hard sphere is given by \begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
The mayer function can be written in terms of weight functions if two 
additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}  where the vector weights are defined as \begin{equation}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
\begin{equation}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation}

@@

Incorporating into Equation~\ref{FexfromSPT} the two vector weights which 
were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated 
expression for $\Phi_3$ developed by Tarazona~\cite{tarazonaphi3, santos2012phi3}, 
the excess free energy becomes
\begin{displaymath}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}}\color{white}\right)}\end{displaymath}
\begin{equation}{\color{white}\left(\color{black}+\frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}\right)\color{black}V}\end{equation} 
where the tensor weight is given by
\begin{equation}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
%and $\Phi_3$ is now
and
\begin{align}
\Phi_1 &= -n_{0}ln(1-n_{3}) \\
\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  
\end{align}

The expression for the weighted densities can be written in a way that 
accomodates both homogenious fluids where the number denstiy n is constant, 
and inhomogeneous fluids where the number density becomes a function of 
the position $n(\vec{r})$.
\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
%The equatons above were derived for a uniform fluid where the number density n is constant, 
%but they can be extended for inhomogenious fluids where the number density becomes a function of r, $n(\vec{r})$.
For a spatially varying number density $n(\vec{r})$, the excess Free 
Energy becomes a functional of the number density profile $n(\vec{r})$ :
\begin{equation}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation}

@@

Our goal is to generate computationally temperature-density and pressure-
temperature phase diagrams for a Weeks-Chandler-Anderson model fluid using 
classical Density Functional Theory. 
A temperature-density phase diagram for a fluid that does not
have attraction between the particles that make up the fluid will show
two discinct regions - one at lower densities, and one at higher densities,
seperated by a coexistence region in which there is found a mixture of the 
fluid at some fixed lower density, and at some fixed higher density. An 
abrupt change from some maximum lower density to some minimum higher density 
occurs for certian combinations of pressure and temperature which form
a line on a pressure-temperature phase diagram. This line distinguishes 
the region in which the fluid exhibits lower densities from the region 
in which the fluid exhbits higher densities. These two regions are known
as the liquid state and the solid state of the fluid.

It is known that a fluid with a soft potential like the WCA fluid will 
form a Face Centered Cubic (FCC) crystal in its solid state~\cite{Hansen}. 
The transition between the liquid 
and solid states is abruptly inhomogeneous, and we study the behavior 
of the fluid by treating the liquid as homogeneous, and by constructing 
a crystal lattice in real space for the solid. Although the entire 
crystal lattice can be constructed with FCC cells, 
FCC cells are not the smallest cells with which the lattice can be built. 
The shaded parallepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
is the smallest structure that, when repeated, re-creates the entire crystal 
lattice. This parallepiped primitive cell takes up one-fourth the volume 
of an FCC cell, and contains the equivalent of one atom.
It includes portions of eight atoms each located on a lattice site convenietly 
reachable by moving along one of three different lattice vectors to step 
from one atom to the next. The three lattice vectors will be useful in 
creating the lattice structure computationally, and will also make the 
program run faster.

@@

It may be that not all lattice sites are occupied. Vacancies can be taken 
into account while keeping %the volume of the bold box, and thus 
the overall number density the same. This is done by 
making the cells smaller. A large collection of smaller cells creates a 
new number density profile shown in Figure \ref{fig:Ensemble_Smallcells} 
with eight smaller Gaussians of one-eigth the height. 
Instead of thinking of an atom in one of the eight small cells, the same 
Gaussian distribution can be obtained from thinking of one-eigth of an 
atom in each small cell. This makes it convenient to describe the new 
number density in terms of the fraction of vacancies, $f_v$: 
\begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath} 
%\begin{align}
  %n(\vec{r})=(1-f_v)\frac{\exp^{-\frac{{\vec{r}}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}
%\end{align}

@@

%\section{Program Flowchart}

>>>>

The goal of this project is to generate temperature-density 
and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
model fluid using classical Density Functional Theory (cDFT). 
To do this, the Helmholtz free energies of the fluid
in a liquid state and in a crystalline solid state must be found.
In general, the Helmholtz free energy, which consists of an ideal part 
and an excess part, can be written as a functional of the number density 
profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}

Soft Fundamental Measure Theory (SFMT) will be used to form the 
functional for the excess Helmholtz free energy. 
SFMT is a method that makes the task of 
forming an expression for the excess Helmholtz free energy of a particular 
model fluid much easier since the interaction that each sphere, or atom, 
has with every other sphere does not have to be taken explicitly into account.
Once the functional for the Helmholtz free energy is formed, classical 
Density Functional Theory is used to find the equilibrium 
Helmholtz free energy at a particular temperature and density. 
According to cDFT the equilbrium 
Helmholtz free energy of the system can be found by varying the number 
density profile $n(\vec{r})$ until the free energy is minimized. 
Results will be compared to those from Monte-Carlo simulations.

@@

Real fluids and their behavior are very complex from a physicist's point 
of view. Although a water molecule, for example, is composed of only three
atoms, the ever-changing redistribution of charges within the molecules, 
and their interaction with charges from other molecules make it difficult 
to study a fluid composed of water molecules directly. The task becomes 
even more daunting considering the complicated molecular motion of each 
water molecule, and the sheer number of molecules that must be taken 
into account. 
The problem of understanding the behavoir of real liquids is better addressed 
by using models which reduce a complex picture to a simple one. Modifications, 
or perturbations, can then be incorporated into the model making it 
progressively more complex until it mimics the behavior of a real 
liquid as closely as possible.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{plot_HardSphere_Potential.pdf}
    \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity 
    for $r\leq{2R}$ where R is the radius of a hard sphere (which is 
    approximately 1.225/2 in this example).}
    \label{fig:HardSphere_potential}
  \end{figure}

@@

The repulsive portion of the Leonard-Jones potential forms the potential for a 
Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship}, 
shown in Figure \ref{fig:WCA_potential}. 
The WCA potential is described by the equation: 
\begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation} 
As seen in Figure \ref{fig:WCA_potential} the WCA potential decreases 
from infinity as the distance between the centers of the atoms r increases, 
and becomes, and remains, essentially zero when 
r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
where R is the radius of the sphere.

As with the Lenard-Jones potential, the repulsive WCA potential increases 
rapidly and approches infinity as the two spheres begin to overlap. This 
is a more realistic repulsive potential than that of the Hard Sphere model 
which goes instantly to infinity when the two spheres make contact (at r=2R), 
and is zero otherwise. Because the WCA spheres have no hard boundary in 
their interatomic potential, they are called ``soft spheres'' and their 
gradually fading potential makes them slightly ``squishy'', as illustrated in
Figure~\ref{fig:TwoSpheres}.

@@

%For each particular temperature, the pressure of the liquid and the 
%solid are the same at the point of transistion. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{P-T_Diagram.png}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
  \end{figure}
The pressure can be related to the Helmholtz free energy F, which is minimized 
for a system at fixed temperature, volume and number of particles as the 
entropy of the system and its surroundings is maximized. The negative of 
the partial derivative of the Helmholtz free energy with respect to volume V 
at a fixed temperature T and number of particles N equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the 
number of particles, the partial derivative of the Helmholtz free energy 
with respect to volume becomes the partial derivative of the Helmholtz 
free energy per atom, $f$,  with respect to the inverse number density. 
\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.

%In a process called Maxwell's Double Tangent construction, two curves are 
%used to find the densities of the liquid and the solid at the point of transistion. 
%Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for 
%the solid both at a fixed temperature. Since the pressure of the liquid 
%is the same as the pressure of the solid at the point of transistion, the 
%density of the liquid and the density of the solid at the point of freezing 
%can be found from identifying the inverse densties at which the tangents 
%to each curve have the same slope.%
Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
and one for the solid at a fixed temperature. The slope of the curves is 
related to the pressure. Since the pressure of the 
liquid is the same as the pressure of the solid at the point of transistion, 
the density of the liquid and the density of the solid at the point of 
freezing can be found from identifying the inverse densties at which the 
tangents to each curve have the same slope. This process is called 
Maxwell's Double Tangent Construction.

@@

%\begin{displaymath}{F=-PV+G}\end{displaymath} 
\begin{align}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{align} 
\begin{align}{f=-P\frac{1}{n}+g}\end{align}
This equation has the form of a line:
\begin{align}{y=mx+y_o}\end{align}
%\begin{displaymath}{G_1=G_2}\end{displaymath} 
%\begin{displaymath}{F_1+PV_1=F_2+PV_2}\end{displaymath} 
%\begin{displaymath}{\frac{F_1}{N}+P\frac{V_1}{N}=\frac{F_2}{N}+P\frac{V_2}{N}}\end{displaymath} 
and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
pressure given by the negative of the slope of the line, and the Gibb's 
free energy per atom given by the y-intercept. 
At the point of phase transition the chemical potential of the liquid is 
the same as the chemical potential of the solid.
\begin{align}{g_L=g_S}\end{align} 
\begin{align}{f_L+P\frac{1}{n_L}=f_S+P\frac{1}{n_S}}\end{align}

The pressure at the transistion point can be obtained from the point of 
intersection that occurs when the Gibbs free energy per atom is plotted 
against the pressure for both the liquid and the solid at fixed temperature, 
as shown in Figure~\ref{fig:GibbsvsP}. These curves are constructed from 
the slope and cooresponding y-intercept data collected from lines tangent 
along the curves on the the Helmholtz free energy per atom verses inverse 
density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
liquid and the solid at a given temperature. 
\begin{figure}[h!]
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
     %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

@@

\begin{displaymath}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{displaymath}    \begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$. 
%\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    
%$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%  
The function $f_{ij}$ is called a Mayer function, and is given by
\begin{equation}\label{eq:mayerfunction}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
Multiplying the Mayer function terms, the excess free energy can be written as
\begin{equation}\label{eq:Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)  }}\end{equation}

@@

The weighted densities can now be expressed as
\begin{equation}{\color{red} CHECK!~~\color{black}n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}}\end{equation} 
with weight functions
\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}

%\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{R}}}\end{equation}
\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
%\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{R}^2}}\end{equation}
\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
where the weight functions $w_{2}$ and $w_{3}$ have the relation
\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}

The weight functions can be related to the mayer function 
\begin{equation}{f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1}\end{equation} 
where r is the distance between the centers of two hard spheres. 
For a hard sphere fluid the interatomic potential $V(r)$ is given by 
\begin{displaymath}{V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.}\end{displaymath}
and so the mayer function for a hard sphere is given by 
\begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
The mayer function can be written in terms of weight functions if two 
additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}  
where the vector weights are defined as 
\begin{equation}\label{eq:w_v2}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
\begin{equation}\label{eq:w_v1}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation}

@@

Incorporating into Equation~\ref{FexfromSPT} the two vector weights which 
were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated 
expression for $\Phi_3$ developed by Tarazona~\cite{tarazonaphi3, santos2012phi3}, 
the excess free energy becomes
\begin{displaymath}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}}\color{white}\right)}\end{displaymath}
\begin{equation}{\color{white}\left(\color{black}+\frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}\right)\color{black}V}\end{equation} 
where the tensor weight is given by
\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
%and $\Phi_3$ is now
and
\begin{align}
\Phi_1 &= -n_{0}ln(1-n_{3}) \label{eq:Phi1}\\
\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \label{eq:Phi2} \\
\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
\end{align}

The expression for the weighted densities can be written in a way that 
accomodates both homogenious fluids where the number denstiy n is constant, 
and inhomogeneous fluids where the number density becomes a function of 
the position $n(\vec{r})$.
\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
%The equatons above were derived for a uniform fluid where the number density n is constant, 
%but they can be extended for inhomogenious fluids where the number density becomes a function of r, $n(\vec{r})$.
For a spatially varying number density $n(\vec{r})$, the excess Free 
Energy becomes a functional of the number density profile $n(\vec{r})$ :
\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation}

@@

Our goal is to generate temperature-density and pressure-temperature 
phase diagrams for a Weeks-Chandler-Anderson model fluid using 
classical Density Functional Theory. 
A temperature-density phase diagram for a fluid that does not
have attraction between the particles that make up the fluid will show
two discinct regions - one at lower densities, and one at higher densities,
seperated by a coexistence region in which there is found a mixture of the 
fluid at some fixed lower density, and at some fixed higher density. An 
abrupt change from some maximum lower density to some minimum higher density 
occurs for certain combinations of pressure and temperature which form
a line on a pressure-temperature phase diagram. This line distinguishes 
the liquid region which exhibits lower densities from the 
solid which exhbits higher densities.

The transition between the liquid 
and solid state is abruptly inhomogeneous in terms of the number density, 
and we study the behavior 
of the fluid by treating the liquid as having a homogeneous number density, 
and by constructing a crystal lattice in real space for the solid. 
It is known that a fluid with a soft potential like the WCA fluid will 
form a Face Centered Cubic (FCC) crystal in its solid state~\cite{Hansen}. 
Although the entire 
crystal lattice can be constructed with FCC cells, 
FCC cells are not the smallest cells with which the lattice can be built. 
The shaded parallepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
is the smallest structure that, when repeated, re-creates the entire crystal 
lattice. This parallepiped primitive cell takes up one-fourth the volume 
of an FCC cell, and contains the equivalent of one atom.
It includes portions of eight atoms each located on a lattice site convenietly 
reachable by moving along one of three different lattice vectors to step 
from one atom to the next. The three lattice vectors will be useful in 
creating the lattice structure computationally, and will also make the 
program run faster.

@@

It may be that not all lattice sites are occupied. Vacancies can be taken 
into account while keeping %the volume of the bold box, and thus 
the overall number density the same by 
making the cells smaller. A large collection of smaller cells creates a 
new number density profile shown in Figure~\ref{fig:Ensemble_Smallcells} 
with eight smaller Gaussians each of which is now one-eigth the height. 
Instead of thinking of an atom in one of the eight small cells, the same 
Gaussian distribution can be obtained from thinking of one-eighth of an 
atom in each small cell as shown in Figure~\ref{fig:SameStatPic}. 
This makes it convenient to describe the new 
number density in terms of the fraction of vacancies, $f_v$: 
\begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath} 
%\begin{align}
  %n(\vec{r})=(1-f_v)\frac{\exp^{-\frac{{\vec{r}}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}
%\end{align}

@@

\section{Program Outline}
